---
title: "Guided Project: Investigating COVID-19 Virus Trends"
author: "Mobin Anandwala"
date: "July 17, 2020"
output: html_notebook
---

# Introduction

This project will be looking at COVID-19 virus trends.  Dataset is from [here](https://www.kaggle.com/lin0li/covid19testing) gathered from January 20th 2020 to June 1st 2020.

# Understanding the Data

We will now load the `tidyverse` library which contains the packages we need to load and understand the data
```{r eval=FALSE}
library(tidyverse)
```


```{r}
covid_df <- read_csv("covid19.csv")
```

```{r}
# Get dimensions names
dim(covid_df)

# Get column names
vector_cols <- colnames(covid_df)

# Display vector columns
vector_cols
```

The dataset contains `10903` rows and `14` columns.  The `vector_cols` vector is a `r typeof(vector_cols)` vector

```{r}
# Display the first few rows
head(covid_df)
```

```{r}
# Get summary of the dataframe using glimpse
glimpse(covid_df)
```

glimpse does the same thing as the first three functions

# Isolating the rows we need
We will now filter the `Province_State` column to keep only the value `"All States"` this way we have country only data and then remove the column `Province_state`
```{r}
covid_df_all_states <- covid_df %>% filter(Province_State == "All States") %>% select(-Province_State)
```

```{r}
glimpse(covid_df_all_states)
```

We can remove the `Province_State` column because we have values for all states here

# Isolating the columns we need
Here the original columns from the dataset

1. `Date`: Date
2. `Continent_Name`: Continent names
3. `Two_Letter_Country_Code`: Country codes
4. `Country_Region`: Country names
5. `Province_State`: States/province names; value is `All States` when state/provincial level data is not available
6. `positive`: Cumulative number of positive cases reported.
7. `active`: Number of actively cases on that **day**.
8. `hospitalized`: Cumulative number of hospitalized cases reported.
9. `hospitalizedCurr`: Number of actively hospitalized cases on that **day**.
10. `recovered`: Cumulative number of recovered cases reported.
11. `death`: Cumulative number of deaths reported.
12. `total_tested`: Cumulative number of tests conducted.
13. `daily_tested`: Number of tests conducted on the **day**; if daily data is unavailable, daily tested is averaged across number of days in between.
14. `daily_positive`: Number of positive cases reported on the **day**; if daily data is unavailable, daily positive is averaged across number of days in.


We want to select columns that show the daily occurance
```{r}
covid_df_all_states_daily <- covid_df_all_states %>% select(Date, Country_Region, active, hospitalizedCurr, daily_tested, daily_positive)

head(covid_df_all_states_daily)
```

# Extracting the top ten tested cases by country

```{r}
covid_df_all_states_daily_sum <- covid_df_all_states_daily %>% group_by(Country_Region) %>% 
  summarise(tested = sum(daily_tested),
            positive = sum(daily_positive),
            active = sum(active),
            hospitalized = sum(hospitalizedCurr)) %>%
  arrange(desc(tested))

covid_df_all_states_daily_sum
```

```{r}
# Taking the top 10
covid_top_10 <- head(covid_df_all_states_daily_sum, 10)

covid_top_10
```

# Identifying the Highest Positive Against Tested Cases

We will now identify the highest positve cases compared to tested cases using the `covid_top_10` dataframe

```{r}
countries <- covid_top_10$Country_Region
tested_cases <- covid_top_10$tested
positive_cases <- covid_top_10$positive
active_cases <- covid_top_10$active
hospitalized_cases <- covid_top_10$hospitalized
```

Assigning country names to each of the cases
```{r}
names(tested_cases) <- countries
names(positive_cases) <- countries
names(active_cases) <- countries
names(hospitalized_cases) <- countries
```

We'll now look at the `positive_cases` vector
```{r}
positive_cases
```

```{r}
# statistics for positive_cases
sum(positive_cases) # total
mean(positive_cases) # average
percent_positive_cases <- positive_cases/sum(positive_cases)
percent_positive_cases
```

Seeing positive cases as a percentage of tested cases
```{r}
positive_cases/tested_cases
```

```{r}
(positive_cases/tested_cases) * 100
```

The top three are the United Kingdom, United States and Turkey (decimals are from rounding down)
```{r}
positive_tested_top_3 <- c("United Kingdom" = 0.11, "United States" = 0.10, "Turkey" = 0.08)
positive_tested_top_3
```

# Keeping relevant information
Now that we have the top 3 countries we don't want to lose the other information that we had
```{r}
united_kingdom <- c(0.11, 1473672, 166909, 0, 0)
united_states <- c(0.10, 17282363, 1877179, 0 ,0)
turkey <- c(0.08, 2031192, 163941, 2980960, 0)

# Create the matrix
covid_mat <- rbind(united_kingdom, united_states, turkey)

# Rename columns
colnames(covid_mat) <- c("Ratio", "tested", "positive", "active", "hospitalized")

covid_mat
```

# Putting it all together

Now we will answer the following question using our matrix and vectors and also create lists of everything that we have

The question is Which countries have had the highest number of positive cases against the number of tests?
```{r}
question <- "Which countries have had the highest number of positive cases against the number of tests?"
answer <- c("Positive tested cases" = positive_tested_top_3)
```

Generating our lists of 

* dataframes
* Matrix
* vectors

```{r}
data <- list(original = covid_df,
             all_states = covid_df_all_states,
             daily = covid_df_all_states_daily,
             top_10 = covid_top_10)

matrices <- list(covid_mat)

vectors <- list(vector_cols, countries)

data_structures_list <- list("dataframes" = data, "matrix" = matrices, "vectors" = vectors)

covid_analysis_list <- list(question, answer, data_structures_list)

covid_analysis_list[[2]]
```

