---
title: "Guided Project: Predicting Condominium Sale Prices"
author: "Mobin Anandwala"
date: "July 14 2020"
output: html_notebook
---

# How well does the size of a Condominium in New York City Explain Sale Price?

This project will explore the following questions:

1. How well does the size of a condominium (measured in gross square feet) explain or predict sale price across New York City as a whole? 
2. How well does the size of a condominium explain or predict sale price for each individual borough?

To answer question 1 we will explore property data for all five boroughs

To answer question 2 we will build, analyze and compare linear models for each borough

# Understanding the data

The data that we are using from this project come from five separate excel files (one for each borough) and can be obtained from [here](https://www1.nyc.gov/site/finance/taxes/property-rolling-sales-data.page)

The data is for June 2019-May 2020

```{r}
library(readxl)
library(readr)
library(stringr)
library(dplyr)
library(tidyr)
library(magrittr)
library(ggplot2)
```

```{r}
# loading in the data from each excel file setting skip = 4 to avoid header rows
queens <- read_excel("rollingsales_queens.xls", skip = 4)
staten_island <- read_excel("rollingsales_statenisland.xls", skip = 4)
brooklyn <- read_excel("rollingsales_brooklyn.xls", skip = 4)
bronx <- read_excel("rollingsales_bronx.xls", skip = 4)
manhattan <- read_excel("rollingsales_manhattan.xls", skip = 4)
```

Each Borough in New York has a specific number which we can get from the `BOROUGH` variable in each dataframe

```{r}
head(queens$BOROUGH, 1)
head(staten_island$BOROUGH, 1)
head(brooklyn$BOROUGH, 1)
head(bronx$BOROUGH, 1)
head(manhattan$BOROUGH, 1)
```

```{r}
# The Borough numbers are as follows:
# Manhattan = 1
# Bronx = 2
# Brooklyn = 3
# Queens = 4
# Staten Island = 5
```

We will now bind the five dataframes into a single dataframe
```{r}
NYC_property_sales <- bind_rows(manhattan, bronx, brooklyn, queens, staten_island)
```

Since we have a single dataframe with all NYC property sales, we don't need the 5 individual dataframes and can remove then using the `rm` function
```{r}
rm(brooklyn, bronx, manhattan, queens, staten_island)
```

Since we have the numbers for each borough we can replace them with the borough names so that it is more clear
```{r}
NYC_property_sales <- NYC_property_sales %>% mutate(BOROUGH = 
                                                      case_when(BOROUGH == 1 ~ "Manhattan",
                                                                BOROUGH == 2 ~ "Bronx",
                                                                BOROUGH == 3 ~ "Brooklyn",
                                                                BOROUGH == 4 ~ "Queens",
                                                                BOROUGH == 5 ~ "Staten Island"))

```

Now that we have our data in a single dataframe with the `BOROUGH` column taken care of there are a few more steps to get this into the format that we need
```{r}
# Convert all column names to lowercase and replace spaces with underscores
colnames(NYC_property_sales) %<>% str_replace_all("\\s", "_") %>% tolower()

# Convert CAPTIALIZED case to Title case (the NEIGHBORHOOD,BUILDING CLASS CATEGORY and ADDRESS columns)
NYC_property_sales <- NYC_property_sales %>% mutate(neighborhood = str_to_title(neighborhood)) %>% mutate(building_class_category = str_to_title(building_class_category)) %>% mutate(address = str_to_title(address))
```

```{r}
# Drop the ease-ment column as it has no data and only select distinct values
NYC_property_sales <- NYC_property_sales %>% select(-`ease-ment`) %>% distinct()

# Filter out property exchanges between family members with a threshold of 10,000
NYC_property_sales <- NYC_property_sales %>% filter(sale_price > 10000) %>%
# Remove properties that have gross_square_feet of 0 
filter(gross_square_feet > 0) %>%
# Remove NA values in both sale_price and gross_square_feet
drop_na(c(gross_square_feet, sale_price))

# Arrange by borough and neighborhood
NYC_property_sales <- NYC_property_sales %>% arrange(borough, neighborhood)

```

```{r eval=FALSE}
# Save the dataframe to a csv file
write_csv(NYC_property_sales, "NYC_property_sales.csv")
```

We now have a csv file that we can use from the original datasets

# Explore Bivarate Relationships with Scatterplots

We will focus our analysis on condominiums with elevators which according to this [document](https://www1.nyc.gov/assets/finance/jump/hlpbldgcode.html) are classified as R4

```{r}
NYC_Condos <- NYC_property_sales %>% filter(building_class_at_time_of_sale == "R4")
```

```{r}
head(NYC_Condos)
```


We will generate a scatterplot using `sale_price` and `gross_square_feet` as those columns are what we are looking at to answer our questions
```{r}
ggplot(data = NYC_Condos,
       aes(x = gross_square_feet, y = sale_price)) +
  geom_point(aes(color = borough), alpha = 0.3) +
  scale_y_continuous(labels = scales::comma, limits = c(0, 75000000)) +
  xlim(0, 10000) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_minimal() +
  labs(title = "Condominium Sale Prices in NYC Generally Increases with Size",
       x = "Size (Gross Square Feet)",
       y = "Sale Price (USD)")
```

```{r}
ggplot(data = NYC_Condos,
       aes(x = gross_square_feet, y = sale_price)) +
  geom_point(aes(color = borough), alpha = 0.3) +
  scale_y_continuous(labels = scales::comma, limits = c(0, 20000000)) +
  xlim(0, 5000) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_minimal() +
  labs(title = "Condominium Sale Prices in NYC Generally Increases with Size",
       x = "Size (Gross Square Feet)",
       y = "Sale Price (USD)")
```

We will use x and y scales specific to each borough and neighborhood
```{r}
ggplot(data = NYC_Condos,
       aes(x = gross_square_feet, y = sale_price)) +
  geom_point(alpha = 0.3) +
  facet_wrap(~ borough, scales = "free", ncol = 2) +
  scale_y_continuous(labels = scales::comma) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_minimal() +
  labs(title = "Condominium Sale Prices in NYC Generally Increases with Size",
       x = "Size (Gross Square Feet)",
       y = "Sale Price (USD)")
```

Somehow the gross_square_feet went missing in data cleaning for condos which is weird as this is the latest data

# Outliers and Data Integrity Issues
```{r}
# Get the highest to lowest sale price
NYC_Condos %>% arrange(desc(sale_price)) %>% head()
```

The highest priced listing was 220 Cenral Park South Ph3 with a square footage of 9,817 (data is found from [here](https://www.cityrealty.com/nyc/MidtownWest/-220-centralparksouth/apartment-PH73/sold/A274304))

```{r}
# Making a copy
NYC_Condos_Original <- NYC_Condos
```

```{r}
# Removing 165 East 66th street
NYC_Condos <- NYC_Condos %>% filter(address != "165 East 66th St, Resi")
```

Let's look at the highest sale in Brooklyn
```{r}
brooklyn_condos <- NYC_Condos %>% filter(borough == "Brooklyn") %>% arrange(desc(sale_price))
head(brooklyn_condos)
```

There is a single unit with a sales price of \$20,301,000 in the Brooklyn Heights neighborhood

Let's get the multiunit sales 
```{r}
multi_unit_sales <- NYC_Condos %>% group_by(sale_price, sale_date) %>% filter(n() >= 3) %>% arrange(desc(sale_price))
```

We will then drop those values in `NYC_Condos` that match in `multi_unit_sales`

```{r}
NYC_Condos <- NYC_Condos %>% anti_join(multi_unit_sales)
```

# Linear Regression Model for Boroughs in New York City Combined

We will now generate the linear models
```{r}
NYC_condos_lm <- lm(sale_price ~ gross_square_feet, data = NYC_Condos)
summary(NYC_condos_lm)
```

