---
title: "Blackhawks vs Oilers Qualifying Round 2020"
author: "Mobin Anandwala"
date: "August 10, 2020"
output: 
  md_document:
    variant: markdown_github
---

# Introduction

We will look at the results of the 2020 Qualifying Round between the Chicago Blackhawks and the Edminton Oilers

This analysis will look at the following:

1. Goals scored overall
2. How the captians (Connor McDavid and Jonathon Toews) Performed goal wise

# Importing the Data

```{r message=FALSE}
library(tidyverse)
```

We will run a scraper that will pull in the data we need (this is already included in the `hockey_scraper.R` script that I built earlier)
```{r}
source('hockey_scraper.R')
```

We will clean up the workspace as we now have our data in the csv file

```{r eval=FALSE}
rm(list = ls())
```

```{r}
CHI_Playoff_data <- read_csv("CHI_Playoff_data.csv")
```

# Data Cleaning

Our analysis is only concerned with the following how many goals were scored overall throughout the series and the performance of the two captains

Performance in this case will be defined as goals and assists

```{r}
goals <- CHI_Playoff_data %>% filter(event_type == "GOAL")
```

For now we will leave the player names as they are simply becase they do not distract from the analysis we are performing (they are all captial)

# Team Analysis

Now we will look at the goals per game
```{r}
goals <- goals %>% mutate(goals = ifelse(event_type == "GOAL", 1, 0))
```

```{r}
goals_per_game <- goals %>% group_by(game_id) %>% summarise(total_goals = sum(goals))
```
Let's make this a little bit easier to understand as we want to see a team breakdown per game

```{r}
goals_per_game_team <- goals %>% filter(!is.na(event_team)) %>% group_by(game_id, event_team) %>% summarise(total_goals = sum(goals))
```


Plotting goals per game between two teams
```{r}
ggplot(data = goals_per_game_team) +
  geom_col(aes(x = game_id, y = total_goals, fill = event_team)) +
  labs(title = "Blackhawks vs Oilers Performance in 2020 Qualifiers", y = "Total Goals per game")
```

# Captains performance

We have so far looked at the overall team performance now let's look at the captains compared to each other in this series

The series was a best of 5 meaning 3 games are needed to advance unlike a best of 7 series where 4 games are needed to advance

```{r}
captain_goals <- goals %>% filter(event_player_1 %in% c("CONNOR.MCDAVID", "JONATHAN.TOEWS"))
```

Note that I am **not** looking at how the captains did point wise which also includes assists just how many goals these two players scored

Applying cleaning so that we can perform analysis
```{r}
captain_goals <- captain_goals %>% mutate(goals = ifelse(event_type == "GOAL", 1, 0))
```

Finally applying split-apply-combine using `group_by` and `summarise`
```{r}
captain_performance <- captain_goals %>% group_by(event_player_1, game_id) %>% summarise(total_captain_goals = sum(goals))
head(captain_performance)
```

Finally plotting the result (the hat trick that McDavid had is giving him the edge here)
```{r}
ggplot(data = captain_performance) +
  geom_col(aes(x = game_id, y = total_captain_goals, fill = event_player_1)) +
  labs(title = "Goal Performance of Captains During Qualifiers", x = "Game ID", y = "Total Goals Scored")
```

